
Protocol
1. Open the R console and load the ggplot2 package
Launch an R or RStudio console session. Load ggplot2 with the following command.

```{r}
library(ggplot2)
library(tidyverse)
```

2. Import data
Read in the data with the following command (you may need to change your working directory or provide the path to where you stored this data file):

```{r}
mutationRate = read.delim ("MutationRate.txt")
```

Check out the structure of the dataset by using the str functioning and printing the first five rows of data with the following commands:

```{r}
str(mutationRate)

mutationRate[1:5,]
```

Create plot.

```{r}
p <- ggplot(data = mutationRate, aes(x = Genotype, y = MutationRate, color = Genome)) +
geom_jitter(width = .2, size = 2,alpha = .5)+
  scale_y_log10()+
  facet_grid(Genome ~ MutationType)

p

# p <- ggplot(data = mutationRate, aes(x = Genotype, y = MutationRate, color = Genome)) +
# geom_point(width = .2, size = 2,alpha = .5)+
#   scale_y_log10()+
#   facet_grid(Genome ~ MutationType)
# 
# p
```

I chose to plot Genotype on X axis and Mutation Rate on Y axis because that is the basis of the biological question. I chose jitter for the geometry so that we could show ALL the data, seeing as there's only three replicates per condition. I chose to use facet_grid to break the plots into 4 combinations of Genome and MutationType variables so that those two other variables could be assessed by the viewer. I redundantly colored the points by Genome because it looks nice. I log transformed the y axis so that the proportional changes of some treatments compared to wild type could more easily be seen. 


There are multiple ways to output a figure to a PDF. One is use the pdf and dev.off commands as follows (you can also explore the ggsave command as well as exporting from the viewer from RStudio).

```{r}
pdf ("mutationRate.pdf")  
p 
dev.off()
dev.off()
```