
```{r}
library(ggplot2)
library(viridis)
library(ggtree)
```



2. Import phylogenetic tree from a newick file
We can use the read.tree function from the ggtree package to read in and store a phylogenetic tree from a file. Our file (tree.nwk) can be downloaded here. This contains the results of a phylogenetic analysis of MutS2 genes from plants and bacteria.

Read in the data with the following command (you may need to change your working directory or provide the path to where you stored this data file):


```{r}
tree = ggtree::read.tree(file = "tree.nwk")
```

Check out the structure of how this tree object is stored in R:

```{r}
tree
```

3. Plot the tree using ggtree
The ggtree packages builds off of ggplot code to visualize phylogenetic trees.


```{r}
ggplot(tree) + geom_tree() + geom_tiplab() + theme_tree()
```


You can use some familiar ggplot commands and parameters to control the visualization.

```{r}
ggplot(tree) + geom_tree(color="firebrick", size=0.5, linetype="dotted", layout="roundrect") + geom_tiplab(size=3) + theme_tree()
```

4. Root the tree
The newick file that we imported contained an unrooted tree, so the visualizations above were arbitrarily rooted. To more accurately represent the history of this gene family, we can set Helicobacter pylori as the outgroup to root the tree.

```{r}
tree_rooted = ape::root(tree, outgroup = "Helicobacter_pylori_AAD07685.1", resolve.root=TRUE)

ggplot(tree_rooted) + geom_tree() + geom_tiplab() + theme_tree()
```
Note that some of our tip lables have been getting cutoff. That is because ggtree sets the default plot size based on the size of the tree without considering the extra label length. We can work around this problem by adding some horizontal expansion to the plot area with hexpand. Alternatively, we could manually set the horizontal dimensions with xlim.

```{r}
ggplot(tree_rooted) + geom_tree() + geom_tiplab() + theme_tree() + hexpand(0.25)
```
You might notice that there is an ancient duplication of this gene such that all green plants and algae represented in this tree have one of each resulting MutS2 gene copy. We can annotate these clades on the tree using geom_hilight and geom_cladelab.

```{r}
ggplot(tree_rooted) + geom_tree() + geom_tiplab(size=3) + theme_tree() + hexpand(0.25) + 
  geom_hilight(node=36, fill="chartreuse4", alpha=.3, type = "gradient", to.bottom=TRUE, extend=1) + 
  geom_hilight(node=44, fill="chartreuse4", alpha=.3, type = "gradient", to.bottom=TRUE, extend=1) + 
  geom_cladelab(node=36, label = "MutS2a", angle=270, hjust = "center", offset=1, offset.text=0.1) + 
  geom_cladelab(node=44, label = "MutS2b", angle=270, hjust = "center", offset=1, offset.text=0.1)
```

5. Read in data associated with the genes represented in the tree
We can read in additional data related to the proteins encoded by the genes in this tree. Our data file (gene_data.txt) can be downloaded here. This file contains information about annotated protein domains in each species as well as in silico predictions of whether the protein is targeted to plastids in plant cells.

```{r}
gene_data = read.delim("gene_data.txt")

gene_data[1:5,]
```

6. Make paneled plots linking associating gene-specifci data with position in the tree
The ggtree package has multiple tools that will add data plots alongside a tree visualization and automatically arrange them such that the data elements are plotted next to the corresponding tip on the tree. The facet_plot command from this package can be used to add plots next to an existing tree, using the typical input parameters and geometries for ggplot2. The following code will generate flanking plots showing protein domain architecture and predicted plastid protein for each sequence in the tree.

```{r}
#store the tree
p1 = ggplot(tree_rooted, aes(x,y)) + geom_tree() + geom_tiplab(size=2.5) + hexpand(1.5) + theme_tree()

#add a facet with lines corresponding to the length of each protein
p2 = facet_plot(p1, panel = "Domains", data = gene_data, geom = geom_segment, mapping = aes(x=0, xend=Seq_Length, y=y, yend=y), size=0.5, color='black')

#add the domains on top of those lines
p3 = facet_plot(p2, panel = "Domains", data = gene_data, geom=geom_segment, aes(x=Domain_Start, xend=Domain_End, y=y, yend=y, col=Domain_Name), size=3) + theme(legend.position = "bottom") + scale_color_viridis(discrete = TRUE)

#add another facet with a bar graph showing the probability of plastid targeting
p4 = facet_plot(p3, panel = "Plastid Targeting", data = gene_data, geom=geom_col, aes(x=Plastid_Targeting), orientation = 'y', fill="chartreuse4") + theme(legend.position = "bottom")

p4
```